A Detailed Analysis of Performance Bottlenecks in MRI Pre-Processing

### 2022-11-10 paper/paper.tex
[ADD] Draft for the introduction.
Plus some more outline for the other sections.

\documentclass[conference,draft]{IEEEtran}

@@

\title{Surveying the Performance Bottleneck of Neuroimaging Pre-Processing}

@@

\section{Introduction}
% TODO

% MRI preprocessing

% Motivate the need to profile the bottleneck of applications.
% * Major part of neuroimaging study
% * Faster pre-processing would enable larger study and more clinical use.
% Identifying the major bottleneck will let us know what to optimize down the line. (compute, compression, etc.)
% Reduced precision + Compression ?

\section{Background}
% TODO
\subsection{MRI Pre-Processing}
% MRI modalities
% Pre-processing tools
% Pre-processing step (fMRIPrep figure, + something for DIPY)

\subsection{Profiling (VTune)}
% Either profiling in general, or VTune in particular.

% VTune
% Lightweight profiler
% Info on function and module runtime

\subsection{Reduced Precision}
% Quickly motivate the need for this paper
% Summarize the advantages and disavantages of using reduced precison.
% Talk about the potential benefits it could have to the field of neurorimaging,
% but that profiling the application hotspot is required to perform it.

\subsection{Compression}
% How it can help with intermediate file.
% Link to reduced precision, which could enable a sort of compression by of it self.

\section{Methods}
\subsection{Infrastructure}
% ??? Should we include that ???
% Doesn't really mattter for this paper, I think.
% But doesn't harm either. For example, the performance could vary depending on architecture.

\subsection{Dataset}
% TODO
% size
% distribution: age, sex, race?, etc.
% Look into CoRR and HPC datasets.
% Need to have quality, yet not prestine, data with subject info for at least age and sex.

\subsection{Pipelines}
% TODO 
% Motivation for those pipelines:

@@

@@

\section*{Data Availability}
% TODO

@@

\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv, paper.bib}

>>>>

\documentclass[conference]{IEEEtran}

@@

\title{A Detailed Analysis of Performance Bottlenecks in MRI Pre-Processing}

@@

\section{Introduction}
Magnetic Resonance Imaging (MRI) is a standard tool used by neuroscientists to
perform clinical diagnosis and for researchers to develop a better understanding
of the brain.
Three main MRI techniques exist: strutural MRI (sMRI), functional MRI (fMRI),
and diffusion MRI (dMRI).
While other modalities such as EEG, CT, and PET exist, we focus on MRIs for
their broad adoption and non-invasive property.
However, MRI data analysis is challeging due to computionally expensive
pre-processing and large output and intermediate data size.

Many toolboxes have been developed to tackle the challenging task of
pre-processing MRI data.
Among those, some openly available and commonly accepted emerged:
AFNI~\cite{Cox1996-sl}, ANTs~\cite{Avants_undated-fu},
FreeSurfer~\cite{Fischl2012-bp}, FSL~\cite{Jenkinson2012-cq}, and
SPM~\cite{Friston2007-ag}.
While the previous pipelines focuse on specific steps of MRI pre-processing,
tools like fMRIPrep~\cite{Esteban2019-og} and DIPY~\cite{Garyfallidis2014-ve}
combine various pipelines to produce an end-to-end pre-processing workflow.

The MRI pre-processing of a subject can take several hours preventing the
clinical use where data analysis might be required in timely manner.
Furthermore, the lenghty pre-processing hinders researcher with limited 
computational resources to perform research on large cohorts.

The first step to improve performance of an application is to understand its
performance bottleneck.
Following, efforts can be focused on those performance critical parts.
For example, better algorithm design or usage of hardware accelerator can 
reduce computation time.
Moreover, compression techniques can be used to lower data transfer time for
output and intermediate data.
Lastly, reduced precision arithmetic can lower computation time,
memory footprint, and storage size.
While these techniques can provide significant performance improvements, their
implementation can require substantial effort.
Therefore, it is critical to understand the performance bottleneck of MRI
pre-processing pipelines to improve their performance.

To the best of our knowledge, there are no comprehensive study of the
bottlenecks in MRI pre-processing.
In this paper, we characterize computational and I/O cost of several commonly
adopted MRI pre-processing pipelines.
The results of our analysis serve as a reference for future efforts to optimise
the MRI pre-processing workflow.

\section{Background}

\subsection{MRI Pre-Processing}
% Discussion of the pre-processing steps.
% Explain their computation and I/O requirements.

% sMRI

% fMRI

% dMRI

\subsection{Intel VTune profiler}
% Requirements for profiling the Neuroimaging field.
% Multi-language support
% Lightweight
% Info on function and module runtime

% VTune

% Challenges remaining
% Debug information when compiling, for valuable information
% Combining the results from multiple analysis. Different input data can 
% vary the analysis significantly. (due to application branching and convergence)

\section{Methods}
\subsection{Infrastructure}
% TODO

\subsection{Dataset}
% TODO
% !!! TBD !!!
% size
% distribution: age, sex, race?, etc.
% Look into CoRR and HPC datasets.
% Need to have quality, yet not prestine, data with subject info for at least age and sex.

\subsection{Pipelines}
% TODO 
% !!! TBD !!!
% Motivation for those pipelines:

@@

\subsection{Profiling}
% How we did the profiling.
% Code to repository
% Aggregation of the results from different subjects.

@@

\section*{Data Availability}
% TODO
The code to compile, profile, and execute the pipelines is publicly available at:

\href{https://github.com/mathdugre/neuro-bottleneck}{https://github.com/mathdugre/neuro-bottleneck}
% TODO Update URL to slashbin org, when created.

@@

\bibliographystyle{IEEEtran}
% \bibliography{IEEEabrv, paper}
\bibliography{paper}
